biclaR: Estimating the socio-environmental impacts of car substitution by bicycle and public transit using open tools

### 2023-10-02 paper/PaperTRA/PaperTRA.Rmd
social vs environmental

abstract: | 
  A high proportion of car trips can be replaced by a combination of public transit and cycling for the first-and-last mile. This paper estimates the potential for cycling combined with public transit (PT) as a substitute for car trips in the Lisbon metropolitan area and assesses its socio-environmental impacts using open data and open source tools.
  A decision support tool that facilitates the design and development of a metropolitan cycling network was developed (_biclaR_). The social and environmental impacts were assessed using the _HEAT for Cycling_ and the _HEAT as a Service_ tools. The impacts of shifting car trips to PT were also estimated and monetized.
  The results indicate that 20% of car trips could switch to the bicycle + PT combination. Shifting to cycling for the first-and-last mile stages can reduce annual CO~2~eq emissions from 6,000 tons/day, with benefits over 10 years of €230 million. For the PT leg, the transfer from car avoids of at least 8,500 tons of CO~2~eq emissions per year.
  This evidence can support policymakers to prioritize interventions that reduce the reliance on private motor vehicles.
keywords: 
  - Active transport
  - Intermodality
  - First and last mile
  - Health economic assessment
  - Environmental impacts
  - Open data and methods
bibliography: bibliography.bib
# Use csl when using default citation_package for Pandoc citeproc
csl: https://www.zotero.org/styles/springer-lecture-notes-in-computer-science
# citations with arabic numbers is prefered when using natbib
natbiboptions: numbers

@@

<!-- TODO: -->
<!-- understand where the values from table 2 come from -->
<!-- check the % of social impacts vs. environmental -->
<!-- make sure of the reference cls -->
<!-- see where we can cut to 1 less paragraph -->

@@

The `r5r` model used the OpenStreetMap road network and the GTFS metropolitan data agregated and validated.
This information is crucial for an accurate PT trip and route estimation.
A digital elevation model from the European Space Agency's COPERNICUS mission, with a 25m spatial resolution, was used to include street gradient information, as a weight in cycling routing.
The cycling potential trips for the two national strategic targets (4% and 10%) was estimated from the values for cycling and car trips (both as a driver and as a passenger) from the 2017 base scenario.

@@

The emissions of CO~2~eq that are avoided during both the initial and final journey segments account for about 75% of the emissions avoided during the PT segment. This finding, while expected -- due the zero cycling emissions, should not be overlooked when promoting the PT use. Improving the safe accessibility to PT interfaces to cyclists and providing bicycle-friendly amenities such as parking facilities can potentially lead to a higher reduction in CO~2~eq emissions, compared to a scenario where individuals shift from car travel to car + PT combination.

The social impacts from replacing car trips to bicycle in first-and-last legs, represent 95% of the total socio-environmental benefits estimated for all journey legs. <!-- sCheck the % of SOCIAL - original heat tables -->
Our findings suggest that a bicycle and PT combination could viably replace 20% of current LMA trips, with an additional 12% of PT journeys prone to further substitution. 
<!-- repeated? -->

>>>>

abstract: | 
  A high proportion of car trips can be replaced by a combination of public transit and cycling for the first-and-last mile. This paper estimates the potential for cycling combined with public transit (PT) as a substitute for car trips in the Lisbon metropolitan area and assesses its socio-environmental impacts using open data and open source tools.
  A decision support tool that facilitates the design and development of a metropolitan cycling network was developed (_biclaR_). The social and environmental impacts were assessed using the _HEAT for Cycling_ and the _HEAT as a Service_ tools. The impacts of shifting car trips to PT were also estimated and monetized.
  The results indicate that 20% of car trips could switch to the bicycle + PT combination. Shifting to cycling for the first-and-last mile stages can reduce annual CO~2~eq emissions from 6,000 tons/day, with benefits over 10 years of €230 million. For the PT leg, the transfer from car avoids of at least 8,500 tons of CO~2~eq emissions per year.
  This evidence can support policymakers to prioritize interventions that reduce the reliance on private motor vehicles.
keywords: 
  - Active transport
  - Intermodality
  - First and last mile
  - Health economic assessment
  - Environmental impacts
  - Open data and methods
bibliography: bibliography.bib
# Use csl when using default citation_package for Pandoc citeproc
csl: https://www.zotero.org/styles/springer-lecture-notes-in-computer-science
# cls : svproc.csl
# citations with arabic numbers is prefered when using natbib
natbiboptions: numbers

@@

<!-- TODO: -->
<!-- understand where the values from table 2 come from -->
<!-- make sure of the reference cls - try the other cls -->
<!-- see where we can cut to 1 less paragraph -->
<!-- should we ignore the social benefits component? and only estimate and add the environmental ones? or estimate the car to tp transfer Social benefits? maybe to the next paper? -->

@@

The `r5r` model used the OpenStreetMap road network and the GTFS metropolitan data aggregated and validated.
This information is crucial for an accurate PT trip and route estimation.
A digital elevation model, from the European Space Agency's COPERNICUS mission, was used to include street gradient information, as a weight in cycling routing. <!--  with a 25 m spatial resolution, -->
The cycling potential trips for the two national strategic targets (4% and 10%) were estimated from the 2017 cycling and car trips (both as a driver and as a passenger), the baseline scenario.

@@

The social impacts represent 98% of the socio-environmental benefits (in value) from replacing car trips to bicycle in first-and-last legs. <!-- We should estimate the social impacts from replacing Car to PT in the second leg, for the other paper -->
The emissions of CO~2~eq that are avoided during both the initial and final journey segments account for about 75% of the emissions avoided during the PT segment. This finding, while expected -- due the zero cycling emissions, should not be overlooked when promoting the PT use.
Improving the safe accessibility to PT interfaces to cyclists and providing bicycle-friendly amenities such as parking facilities can potentially lead to a higher reduction in CO~2~eq emissions, compared to a scenario where individuals shift from car travel to car + PT combination.

Our findings suggest that a bicycle and PT combination could viably replace 20% of current LMA trips, with an additional 12% of PT journeys prone to further substitution. 
<!-- repeated? -->
